%%%%%%%%%%%%%%%%%%%%%%%%%%  ltexpprt.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is ltexpprt.tex, an example file for use with the SIAM LaTeX2E
% Preprint Series macros. It is designed to provide double-column output.
% Please take the time to read the following comments, as they document
% how to use these macros. This file can be composed and printed out for
% use as sample output.

% Any comments or questions regarding these macros should be directed to:
%
%                 Donna Witzleben
%                 SIAM
%                 3600 University City Science Center
%                 Philadelphia, PA 19104-2688
%                 USA
%                 Telephone: (215) 382-9800
%                 Fax: (215) 386-7999
%                 e-mail: witzleben@siam.org


% This file is to be used as an example for style only. It should not be read
% for content.

%%%%%%%%%%%%%%% PLEASE NOTE THE FOLLOWING STYLE RESTRICTIONS %%%%%%%%%%%%%%%

%%  1. There are no new tags.  Existing LaTeX tags have been formatted to match
%%     the Preprint series style.
%%
%%  2. You must use \cite in the text to mark your reference citations and
%%     \bibitem in the listing of references at the end of your chapter. See
%%     the examples in the following file. If you are using BibTeX, please
%%     supply the bst file with the manuscript file.
%%
%%  3. This macro is set up for two levels of headings (\section and
%%     \subsection). The macro will automatically number the headings for you.
%%
%%  5. No running heads are to be used for this volume.
%%
%%  6. Theorems, Lemmas, Definitions, etc. are to be double numbered,
%%     indicating the section and the occurence of that element
%%     within that section. (For example, the first theorem in the second
%%     section would be numbered 2.1. The macro will
%%     automatically do the numbering for you.
%%
%%  7. Figures, equations, and tables must be single-numbered.
%%     Use existing LaTeX tags for these elements.
%%     Numbering will be done automatically.
%%
%%  8. Page numbering is no longer included in this macro.
%%     Pagination will be set by the program committee.
%%
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\documentclass[twoside,leqno,twocolumn]{article}
\usepackage{ltexpprt}
\usepackage{parskip}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage[many]{tcolorbox}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{makecell}
\usepackage{float}
\usepackage{placeins}
\usepackage{microtype}

\newcommand{\Tau}{\mathrm{T}}

\graphicspath{{./images/}}

\begin{document}


%\setcounter{chapter}{2} % If you are doing your chapter as chapter one,
%\setcounter{section}{3} % comment these two lines out.

\title{\Large Clustering Vehicles based on Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
\author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom}
\and
Matthew Forshaw\footnotemark[1]
\and
A. Stephen McGough\footnotemark[1]}
\date{}

\maketitle

% Copyright Statement
% When submitting your final paper to a SIAM proceedings, it is requested that you include
% the appropriate copyright in the footer of the paper.  The copyright added should be
% consistent with the copyright selected on the copyright form submitted with the paper.
% Please note that "20XX" should be changed to the year of the meeting.

% Default Copyright Statement
\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 2018 by BigTraffic\\
Unauthorized reproduction of this article is prohibited}}}

% Depending on which copyright you agree to when you sign the copyright form, the copyright
% can be changed to one of the following after commenting out the default copyright statement
% above.

%\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 20XX\\
%Copyright for this paper is retained by authors}}}

%\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 20XX\\
%Copyright retained by principal author's organization}}}


%\pagenumbering{arabic}
%\setcounter{page}{1}%Leave this line commented out.

\begin{abstract} \small\baselineskip=9pt In major cities, government agencies increasingly employ automatic number-plate recognition (ANPR) technology in law enforcement and traffic control. In the Tyne and Wear region (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads. So far, few works have explored the full potential of number-plate scans for analysing individual and collective travel patterns. In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations. We illustrate the effect of parameters $\tau$ and $\Tau$ on trip discrimination and on the detection of implausible trips. To demonstrate the potential of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over fifteen weekdays. Results show that whilst private and transit travel modes can begin to be inferred from the resulting clusters, further work needs to be put into developing a more consistent and integrated framework for trip identification in ANPR data.\end{abstract}


\input{intro.tex}

\input{dataset.tex}

\input{trips.tex}

\input{classification.tex}

\input{discussion.tex}

\begin{thebibliography}{99}

\bibitem{ODMobileData}
Alexander, Lauren, et al. "Origin–-destination trips by purpose and time of day inferred from mobile phone data." Transportation research part c: emerging technologies 58 (2015): 240-250.

\bibitem{Castillo2010}
Castillo, Enrique, et al. {\em Optimal use of plate-scanning resources for route flow estimation in traffic networks}. IEEE Transactions on Intelligent Transportation Systems 11.2 (2010): 380-391.

\bibitem{Castillo2008}
Castillo, Enrique, José María Menéndez, and Pilar Jiménez. {\em Trip matrix and path flow reconstruction and estimation based on plate scanning and link observations}. Transportation Research Part B: Methodological 42.5 (2008): 455-481.

\bibitem{Clustering}
Chen, Huiyu, Chao Yang, and Xiangdong Xu. {\em Clustering Vehicle Temporal and Spatial Travel Behavior Using License Plate Recognition Data}. Journal of Advanced Transportation 2017 (2017).

\bibitem{noVehiocles}
Department for Transport {\em Provisional Road Traffic Estimates Great Britain: October 2016-September 2017}. url (visited January 2017): https://www.gov.uk\-/government/statistics/provisional-road-traffic-estima\-tes-great-britain-october-2016-to-september-2017.


\bibitem{distance}
Department for Transport {\em Vehicle Licensing Statistics: Quarter 3 (Jul-Sep) 2017}. url (visited 31 January 2017): https://www.gov.uk/government/statistics\-/vehicle-licensing-statistics-july-to-september-2017.

\bibitem{EvolutionUTMC2013}
Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.

\bibitem{FourStepModel}
McNally, Michael G. "The four-step model." Handbook of Transport Modelling: 2nd Edition. Emerald Group Publishing Limited, 2007. 35-53.

\bibitem{Hazelton2012}
Parry, Katharina, and Martin L. Hazelton. {\em Estimation of origin-–destination matrices from link counts and sporadic routing data}. Transportation Research Part B: Methodological 46.1 (2012): 175-188.

\bibitem{ClusteringGPS}
Schüssler, Nadine, and Kay W. Axhausen. {\em Identifying trips and activities and their characteristics from GPS raw data without further information}. Arbeitsbericht Verkehrs-und Raumplanung 502 (2008).

\bibitem{Kmeans}
Steinley, Douglas, and Michael J. Brusco. {\em Choosing the number of clusters in Κ-means clustering}. Psychological Methods 16.3 (2011): 285.

\bibitem{SurveyITS2011}
Zhang, Junping, et al. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.

% \bibitem{ShortestPath?2015}
% Zhu, Shanjiang, and David Levinson. {\em Do people use the shortest path? An empirical test of Wardrop’s first principle}. PloS one 10.8 (2015): e0134322.

% \bibitem{BANKSMITH}
% R.~E. Bank and R.~K. Smith, {\em General sparse elimination requires no
%   permanent integer storage}, SIAM J. Sci. Stat. Comput., 8 (1987),
%   pp.~574--584.
%
% \bibitem{EISENSTAT}
% S.~C. Eisenstat, M.~C. Gursky, M.~Schultz, and A.~Sherman, {\em
%   Algorithms and data structures for sparse symmetric gaussian elimination},
%   SIAM J. Sci. Stat. Comput., 2 (1982), pp.~225--237.

\end{thebibliography}
\end{document}

% End of ltexpprt.tex
